---
title: "Visual Analysis"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    social: menu
    source: embed
    theme: yeti 
---

```{r setup, include = FALSE}
library(flexdashboard)
library(plyr)
library(ggplot2)
library(ggridges)
library(plotly)
library(dplyr)
library(readr)
library(gridExtra)
```

### Top statistical "hot topics" by year from 1888 - 2017

```{r, 1st plot, echo = FALSE, message = FALSE}

jasa_rm_stopwords = read_csv("./analysis/data/jasa_rm_stopwords.csv")
 
# Top word by year, proportion
 
pop_word_prop = jasa_rm_stopwords %>%
  group_by(year) %>%
  dplyr::count(word) %>% # DONT DELETE dplyr::
  mutate(sum = sum(n),
         proportion = n/sum) %>%
  arrange(proportion) %>%
  top_n(1) %>%
  ggplot(aes(x = year, y = proportion, fill = word)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = "Top Words by Year")
 
ggplotly(pop_word_prop)
 
# "statistics" in late 1800's to early 1900's popular
# "data", "models" popular in late 1900's and 2000's ("data" in 2006)
```


- Note: missing years before 1923 are due to two-year chapters, which shorten to one-year chapters following

- "Statistics" and "statistical" dominated word proportion until the 1960s, when words such as "estimation", "models", and "data" began to appear

- Proportions of key words likely decreased over time because as more papers were published into the 1940s, there became more vairability in the titles

### Trends of statistical "hot topics" from 1888 - 2017

```{r, 2nd plot}

words_ridges = jasa_rm_stopwords %>%
  filter(matched_word != "NA") %>%
  ggplot(aes(x = year, y = matched_word, fill = matched_word)) +
  geom_density_ridges() +
  labs(title = "Trends of Hot Topics from 1888 to 2017") + 
  theme(legend.position = "none") +
  ylab("Keyword") +
  xlab("Year")

words_ridges

```

***

- Thirteen key words of interest were chosen to be examined for this plot 

- Words such as "statistic", "sampl", and "estimat" are include all words that begin with such phrases (for example, "statistic" includes "statistical", "statistics", etc.)

- Only "statistic" has remained relatively steady over time

- Large spike in "p-value" after the year 2000, consistent with the growing attention p-values are getting

- Large spike in "bootstrap" during the 1970s-1990s, which has since decreased

- Increase in titles including "data" may be due in part to the growth of large data

### Number of collaborators per article for every year from 1888 - 2017
    
```{r, 3rd plot}

library(janitor)
library(tidyverse)
library(tidytext)
jasa_articles <- read_csv("./analysis/data/jasa_articles_dataset_complete.csv")
 
jasa_articles <- jasa_articles %>%
  clean_names() %>%
  separate(authors, into = c("author1", "author19"), sep = " & ") %>%
  separate(author1, into = paste("author", 1:18), sep = ", ") %>%
  unnest_tokens(word, title)
data("stop_words")
 
jasa_articles = anti_join(jasa_articles, stop_words)
 
jasa_articles <- jasa_articles %>%
  dplyr::select(`author 1`:author19, year)
 
no_authors_year <- jasa_articles %>%
  dplyr::mutate( number_authors = rowSums(!is.na(jasa_articles)) - 1) %>%
  dplyr::select(number_authors, year) %>%
  group_by(year) %>%
  dplyr::summarize( Mean_no_authors = mean(number_authors))
 
finalplot <- no_authors_year %>%
  ggplot(aes(x = year, y = Mean_no_authors, color = Mean_no_authors)) +
  geom_point() +
  geom_smooth(se = FALSE, color = "dodgerblue3") + labs(
    title = "Trend of the mean number of authors",
    x = "Year",
    y = "Mean number of authors"
    ) +
  theme(legend.position = "none")

ggplotly(finalplot)

```
    
***

- "Mean number of authors" corresponds to the mean number of authors cited per paper per year

- There is a clear positive trend in the number of authors cited, illustrating an increase in collaboration on statistical papers

- It will be interesting to note if this trend continues to increase and/or ever reaches a plateau or decrease in the trend
    
### Number of citations per statistical "hot topic" 

```{r, 4th plot}

library(forcats)
library(tidyverse)
jasa_rm_sw = read_csv("./analysis/data/jasa_rm_stopwords.csv")
 
citation_top10 <- jasa_rm_sw %>%
  dplyr::filter(!is.na(matched_word)) %>%
  dplyr::select(id, citations, matched_word) %>%
  group_by(matched_word) %>%
  dplyr::summarize( Total_number_of_citations = sum(citations))  %>%
  rename(word = matched_word)


citation_top10 %>% 
  dplyr::mutate(Hot_topics = fct_reorder(word, Total_number_of_citations)) %>% 
  plot_ly(x = ~Hot_topics, y = ~Total_number_of_citations, color = ~Hot_topics, type = "bar") %>%
  layout(
    xaxis = list(
      type = 'category',
      title = 'Keyword'
    ),
    yaxis = list(
      title = 'Total Number of Citations'
    )
  )
 
```

*** 

- Plot depicts the number of people who have cited each paper with the given statistical "hot topic" in the title

- Words such as "statistic", "sampl", and "estimat" are include all words that begin with such phrases (for example, "statistic" includes "statistical", "statistics", etc.)

- "Estimat", "model", and "regression" generate the greatest number of citations

- "P-value", "clinical", and "bootstrap" generate the fewest number of citations
